Towards Explainability of Machine Learning Models in Insurance Pricing

### 2019-10-24 manuscript/manuscript.Rmd
fix equation alignment

Let us look at some examples of models that are considered by many to be interpretable. The GLM is probably the most oft-cited example of such a model (or rather, modeling framework) in the actuarial literature. Given a set of inputs, we can easily reason about what the output of the model is. As an illustrative example, we may have a claims severity model with driver age, amount of insurance (AOI), and sex as features; assuming a log link function, we have

\begin{equation}
  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{AOI} + \beta_3 \cdot \text{sex}_{\text{male}}.
\end{equation}

@@

\begin{equation}\label{eq:badglm}
  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{AOI}  + \beta_2\cdot\text{age} + \beta_3\cdot\text{age}^2 + \beta_4 \cdot \text{age:AOI} + \beta_5 \cdot \text{gender}_{\text{male}}.
\end{equation}

@@

\begin{equation}\label{eq:var-contrib}
v(j, x_* ) = E_X[\widehat{f}(X) | X^1 = x_* ^1, \dots, X^j = x_* ^j] - E_X[\widehat{f}(X^1 = x_* ^1, \dots, X^{j-1} = x_*^{j - 1})].
\end{equation}

>>>>

Let us look at some examples of models that are considered by many to be interpretable. The GLM is probably the most oft-cited example of such a model (or rather, modeling framework) in the actuarial literature. Given a set of inputs, we can easily reason about what the output of the model is. As an illustrative example, we may have a claims severity model with driver age, sex, and vehicle age as features; assuming a log link function and $Y$ denotes the response, we have

\begin{align}
& \log(E[Y]) = \beta_0 + \beta_1 \cdot \text{age} \nonumber \\
& \quad + \beta_2 \cdot \text{vehicle\_age} + \beta_3 \cdot \text{sex}_{\text{male}}.
\end{align}

@@

\begin{align}\label{eq:badglm}
& \log(E[Y]) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{vehicle\_age} \nonumber\\
& \quad + \beta_3\cdot\text{vehicle\_age}^2 \nonumber \\
& \quad + \beta_4 \cdot \text{age} \cdot \text{vehicle\_age} \nonumber \\
& \quad + \beta_5 \cdot \text{sex}_{\text{male}} + \beta_6 \cdot \text{sex}_{\text{male}} \cdot \text{age}. 
\end{align}

@@

\begin{align}\label{eq:var-contrib}
& v(j, x_* ) = E_X[\widehat{f}(X) | X^1 = x_* ^1, \dots, X^j = x_* ^j] \nonumber\\
& \quad - E_X[\widehat{f}(X^1 = x_* ^1, \dots, X^{j-1} = x_*^{j - 1})].
\end{align}
